### 2017-01-31 zoon_app_cover.Rmd
add cover letter

Dear Professor Freckleton,

I am pleased to submit the article: "The zoon R package for reproducible and shareable species distribution modelling" for consideration as an Application paper.

The Zoon Project was set up to improve reproducibility in the ecological sub discipline of species distribution modeling (SDM). This paper describes the core functionality of the project: the zoon R package, which lets ecologists construct and modify fully-reproducible SDM workflows consisting of community-contributed software modules held in an online repository.

Like the many other SDM R packages, zoon enables users to fit of SDMs with a variety of methods. However the aim and novel contribution of zoon is quite different - to facilitate the SDM community in sharing methods (modules) and experiments (workflows) as standardised, repeatable and reusable artefacts. This aligns with the journal's recent efforts to improve archiving of analytical code and has the same end goal - improving the ecological community's ability to build on each others' work. We therefore believe that our manuscript would be well suited for publication as an Application paper.

Note that two of the authors are MEE associate editors and that both Bob O'Hara and Rich FitzJohn have been involved in the development of Zoon, so one of the three remaining Applications AEs would be best placed to handle the manuscript.

Yours Sincerely,
Nick Golding, on behalf of the authors